```{r echo=FALSE}
yml_content <- yaml::read_yaml("chapterauthors.yml")
author <- yml_content[["lab3shortname"]][["author"]]
```
# Terrain Analysis {#terrain}

Written by
```{r results='asis', echo=FALSE}
cat(author)
```

## Lab Overview {-}

A Digital Elevation Model (DEM) is a digital representation of the Earth's terrain including mountains, valleys, rivers, and other topographic features. They are typically created using remote sensing technology, such as radar or LiDAR (Light Detection and Ranging), which capture elevation data points across the landscape. Typically, these elevation data points are organized into raster format, where each raster cell represents elevation within specific pixel. DEMs are used in a range of applications, including cartography, hydrology, geology, environmental analysis, and simulating water flow and erosion. 

In this lab you will use a DEM and the Hydrology toolset in ArcGIS Pro to map stream networks and watersheds within critical salmon spawning habitat. You will then use data from Climate NA to understand how changes in temperature and precipitation may impact future water flow in these watersheds. 

------------------------------------------------------------------------

### Learning Objectives {-}

- Understand how data is represented in a DEM 

- Learn how to derive slope, aspect and Topograhic Position Index (TPI) using raster focal calculations

- Use the Hydrology Toolbox to map stream networks and watershed boundaries

------------------------------------------------------------------------

### Deliverables {-}

------------------------------------------------------------------------

### Data {-}

- DEM of the XXXX region

------------------------------------------------------------------------

## Task 1: Understanding DEMs {-}

Step 1: Import DEM. What is the resolution of the data? 

What is the highest and lowest elevation in the dataset? 

Step 2: Calculate slope and aspect

Step 3: Topographic Position Index

------------------------------------------------------------------------

## Task 2: Identifying Stream Networks {-}

Step 1: Fill the DEM to remove sinks/holes

Step 2: Calculate Flow Accumulation

How does changing the accumulation threshold change the resulting stream network?  

Step 3: Flow Direction

Step 4: Stream ordering

Step 5: Stream to feature tool to vectorize stream networks

------------------------------------------------------------------------

## Task 3: Mapping Watersheds {-}

Step 1: Snap to pour point

How does changing the location of the pour point impact the results? 

Step 2: Watershed tool

Step 3: Vectorize/visualize watersheds

------------------------------------------------------------------------

## Task 4: Analyze watersheds {-}

Step 1: Import temperature and precipitation trends from Climate NA

Step 2: Summarize climate change variables by watershed

Which watersheds have the greatest change in temperature/precipitation? 

How might these changes impact the stream network you identified in Task 2? 
