```{r echo=FALSE}
yml_content <- yaml::read_yaml("chapterauthors.yml")
author <- yml_content[["lab3shortname"]][["author"]]
```
# Terrain Analysis {#terrain}

Written by
```{r results='asis', echo=FALSE}
cat(author)
```

## Lab Overview {-}

Lorem ipsum

------------------------------------------------------------------------

## Learning Objectives {-}

------------------------------------------------------------------------

## Deliverables {-}

------------------------------------------------------------------------

## Data {-}

DEM

------------------------------------------------------------------------

## Task 1: Preparing the DEM {-}

Step 1: Import DEM. What is the resolution of the data? 

What is the highest and lowest elevation in the dataset? 

Step 2: Fill the DEM to remove sinks/holes

------------------------------------------------------------------------

## Task 2: Identifying Stream Networks {-}

Step 1: Flow Accumulation

How does changing the accumulation threshold change the resulting stream network?  

Step 2: Flow Direction

Step 3: Stream ordering

Step 4: Stream to feature tool to vectorize stream networks

------------------------------------------------------------------------

## Task 3: Mapping Watersheds {-}

Step 1: Snap to pour point

How does changing the location of the pour point impact the results? 

Step 2: Watershed tool

Step 3: Vectorize/visualize watersheds

------------------------------------------------------------------------

## Task 4: Analyze watersheds {-}

Step 1: Import temperature and precipitation trends from Climate NA

Step 2: Summarize climate change variables by watershed

Which watersheds have the greatest change in temperature/precipitation? 

How might these changes impact the stream network you identified in Task 2? 
